---
title: "Term Deposit Classification Problem"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal
 The goal is to predict if a client will subscribe a term deposit.<br> A term deposit is a cash investment held at a financial institution. The money is invested for an agreed rate of interest over a fixed amount of time, or term. It also can be called having a fixed deposit account.
 
## Breakdown
 I will split the process in three. First, i will perform and Exploratory Data Analysis on the dataset now, then build a machine learning model to predict if the client would want to subscribe for a term deposit then build a Web Application for convinence.
 As the data scientist, i would really want to focus on the people that will want to subscribe for a term deposit, knowing the really important factors that influences this will be my end goal. 
 
## Loading Library and Dataset
```{r message=FALSE}
library(tidyverse)


theme_set(ggthemes::theme_gdocs())
```

```{r}
bank_df <- read_delim("bank.csv", ";", escape_double = FALSE, trim_ws = TRUE )

head(bank_df)
```

## Data Dictionary 
<ol>

<li> age (numeric) </li>
<li> job : type of job </li>
<li> marital : marital status </li>
<li> education </li>
<li> default: has credit in default? </li>
<li> balance: average yearly balance, in euros (numeric) </li>
<li> housing: has housing loan? </li>
<li> loan: has personal loan? </li>
   <h5> Related with the last contact of the current campaign: </h5>
<li> contact: contact communication type (categorical: unknown,telephone,cellular) </li>
<li> day: last contact day of the month (numeric) </li>
<li> month: last contact month of year </li>
<li> duration: last contact duration, in seconds (numeric) </li>
   <h5> Other attributes: </h5>
<li> campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact) </li>
<li> pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric, -1 means client was not previously contacted) </li>
<li> previous: number of contacts performed before this campaign and for this client (numeric)  </li>
<li> poutcome: outcome of the previous marketing campaign (categorical: unknown,other,failure,success) </li>
  <h5> Output variable (desired target): </h5>
<li> y - has the client subscribed a term deposit? (binary: "yes","no")</li>
</ol>

## EDA Pipleine
<ul>
<li> Getting a summary Statistics of our dataset </li>
<li> Converting Variables to proper data formats </li>
<li> Check for class Imbalance </li>
<li>Look at the relationship of the predictors to the target outcomes</li>
</ul>

## EDA

```{r}
bank_df <-  bank_df %>% mutate_if(is.character, factor)
```

```{r}
## Smmary Statistics
## Some numeric variables are skewed, Normalization will be required when creating recipe steps
skimr::skim(bank_df)
```

## Test for class imbalance 
```{r}
## From the summary statistics above, this dataset has class imbalance
## To visualize it.
bank_df %>%
   ggplot(aes(y, fill = y))+
   geom_bar(stat = "count", show.legend = F, col = "black", alpha = 0.9)+
   labs(title = "Test for Class Imbalance",
        subtitle = "It is very obvious, there is a serious class imbalance case here",
        x = NULL,
        y = NULL)
```

## Comparing Age to Outcome
```{r}
## I suspect Older people are more likely to subscribe for a term deposit
## This is a dataset about Fixed Deposit accounts, it makes sense not to see younger age groups
bank_df %>%
   ggplot(aes(age, fill = y))+
   geom_histogram()

```


```{r}

bank_df %>%
   mutate(age_groups = cut(age, breaks = seq(18, 90, 10), include.lowest = T, right = F)) %>%
   ggplot(aes(age_groups, fill = y))+
   geom_bar(stat = "count")

## Here we see that age ranges from 29-39 has the highest turnout, but is that so?
## Lets look at their ratio plot
## Interest in Term deposit increases with Age
bank_df %>%
   mutate(age_groups = cut(age, breaks = seq(18, 90, 10), include.lowest = T)) %>%
   ggplot(aes(age_groups, fill = y))+
   geom_bar(stat = "count", position = "fill")+
   scale_y_continuous(labels = scales::percent_format())+
   labs(title = "Relationship between different age groups\n and Opening a Fixed Deposit Account",
        y = "Rate",
        x = "Age Groups",
        fill = "Outcome")


```

## Comparing Job to Outcome
```{r}
## Creating a Positivity Ratio profile on the different occupation inn the dataset.
## The retired people comes first with confirms the question raised before about elderly people more interested in term deposit
## A simple Plot to show our Results
## Missing values are in this variable, "Unknown"
bank_df %>%
   select(job, y) %>%
   group_by(job) %>%
   count(y) %>%
   ungroup() %>%
   mutate(job = str_to_title(job)) %>%
   spread(key= y, value = n) %>%
   mutate(positivity_ratio = round((yes/(yes+no)),3)) %>%
   mutate(job = reorder(job, positivity_ratio)) %>%
   arrange(desc(positivity_ratio)) %>%
   ggplot(aes(positivity_ratio,job, col = job))+
   geom_point()+
   geom_segment(aes(xend = 0, x = positivity_ratio, y = job, yend = job))+
   labs(
      title = "Jobs and their Positivity Rates",
      subtitle = "The Positivity rate here is defined as the percentage of the people\n that subscribed to a term deposit",
      y = NULL,
      x = NULL
   )+
   scale_x_continuous(labels = scales::percent_format())+
   theme(
      legend.position = "none",
      axis.text.y = element_text(face = "bold.italic"),
      plot.subtitle = element_text(face = "italic")
   ) 
 

```

## Relationship between Marital Status and the Outcome.
```{r}
## The Number of Single people in the dataset is twice the number of divorced people
## but 14.6% of the divorced set subscribed to a term deposit.

 bank_df %>%
   select(marital, y) %>%
   group_by(marital) %>%
   count(y)%>%
   spread(key = y, value = n) %>%
   mutate(total = yes+no,
      positivity_ratio = round((yes/total),3)) %>%
   arrange(desc(positivity_ratio)) %>%
   select(marital,total, positivity_ratio) %>%
   mutate(positivity_ratio = paste0(round(100*positivity_ratio,2),"% subscribed to a term deposit"))%>%
   data.frame()

   
```

## Relationship between Education and the Outcome.
```{r}

## There are also missing values here, "unknown".
## Method of Imputation: knnImpute seems to be a better choice.
## We can see that tertiary education category has the highest positivity rate.
bank_df %>%
   select(education, y) %>%
   group_by(education) %>%
   count(y) %>%
   ungroup() %>%
   spread(key = y, value = n) %>%
   mutate(total = no + yes,
          p_rate =round((yes/total)*100,2)) %>%
   select(education,total,p_rate) %>%
   arrange(desc(p_rate))


## Creating a function to produce postivity rate table

p_rate <- function(df,column_name){
   column_name <- enquo(column_name)
   
   df %>%
   select(!!column_name, y) %>%
   group_by(!!column_name) %>%
   count(y) %>%
   ungroup() %>%
   spread(key = y, value = n) %>%
   mutate(total = no + yes,
          p_rate =round((yes/total)*100,2)) %>%
   select(!!column_name,total,p_rate) %>%
   arrange(desc(p_rate))
}
   
p_rate(bank_df,education)

```

## Relationship between Default and the Outcome.
```{r}

## Well, whether a customer has credit in default or not doesn't really seem to affect the outcome
p_rate(bank_df,default)
   
```


## Relationship between Balance and the Outcome.
```{r}
## Does it mean that someone who earns a negative amount is in debt?
bank_df %>%
   select(balance) %>%
   summary()
```

```{r}
## People that earn less than 10,000 Euros tends to subscribe to a term deposit? lets investigate further.
## Normalization is required.
bank_df %>%
   filter(balance > 0) %>%
   select(balance, y) %>%
   ggplot(aes(balance, fill = y)) +
   geom_histogram(show.legend = F)+
   labs(
      x = "Yearly average income"
   )+
   facet_wrap(y~., scales = "free")
```


```{r}
## People that earn above 10,000 Euros yearly, are more likely not to subscribe for a term deposit. 
bank_df %>%
   filter(balance >10000 ) %>%
   select(balance, y) %>%
     ggplot(aes(balance, fill = y)) +
   geom_histogram(show.legend = F)+
   labs(
      x = "Yearly average income"
   )+
   facet_wrap(y~., scales = "free")
```


```{r}
## Also,if one is in debt (negative income), chances are that he/she could subscribe for a term deposit
bank_df %>%
   filter(balance < 0 ) %>%
   select(balance, y) %>%
   ggplot(aes(balance, fill = y)) +
   geom_histogram(show.legend = F)+
   labs(
      x = "Yearly average income"
   )+
   facet_wrap(y~., scales = "free")
```


## Relationship between Housing Loan and the Outcome.
```{r}

## People that don't havea housing loan are almost twice more likely to subscribe for a term deposit.
p_rate(bank_df,housing) 
   





```

## Relationship between Persnal Loan and the Outcome.
```{r}

## People that don't have a personal loan are 2x more likely to subscribe for a term deposit.

p_rate(bank_df,loan)
```

## Relationship between Contact and the Outcome.
```{r}
## There isn't much to differenciate them, although we have some missing values "unknown"

p_rate(bank_df, contact)

```

## Relationship between Days and Months and the Outcome. 
```{r}
## It seems normal
bank_df %>%
   select(day, y) %>%
   ggplot(aes(day, fill = y)) +
   geom_histogram(show.legend = F)+
   labs(
      x = "Last contact day of the month"
   )+
   facet_wrap(y~., scales = "free")
```

```{r}
## thoughts: I would want to combine the month and day to figure out if that day was on a holiday
bank_df %>%
   select(month, y) %>%
   ggplot(aes(month, fill = y)) +
   geom_bar(stat = "count", show.legend = F)+
   labs(
      x = "Last contact month of the year"
   )
```


## Relationship between Call Duration and the Outcome.
```{r}
## I have a crazy thought, i want to know if richer customers had a higher call duration
## Well, there doesn't seem to be any preferencial treatments here.
## People who spent more than 8 minutes on the phone were more likely to subscribe? Lets dig in.

options(scipen = 999)

bank_df %>%
   select(balance, duration, y) %>%
   mutate(duration = duration/60) %>%
   ggplot(aes(y = duration, x = balance, col = y)) +
   geom_point(alpha = 0.3)+
   geom_hline(yintercept = 8)+
   scale_y_log10()+
   scale_x_log10()+
     labs(
      y = "Call Duration in Minutes",
      x = "Yearly average income",
      col = "Outcome"
   ) 


```



```{r}
## People who spent more than 8 minutes on the phone were more likely to subscribe?
## Yes, this is true

bank_df %>%
   mutate(duration = duration/60) %>%
   mutate(test = ifelse(duration > 8,"More than 8 Minutes","Less than 8 Minutes")) %>%
   select(balance, duration, y, test) %>%
   ggplot(aes(y = duration, x = balance, col = y)) +
   geom_point(alpha = 0.6)+
   scale_y_log10()+
   scale_x_log10()+
   facet_wrap(~test, scales = "free")+
   labs(
      y = "Call Duration in Minutes",
      x = "Yearly average income",
      col = "Outcome"
   )+
   ggthemes::theme_fivethirtyeight()
```

## Relationship between Campaign and the Outcome.
```{r}

## People that subscribed didn't need to have too many campaigns
bank_df %>%
   select(campaign,y) %>%
   ggplot(aes(campaign, fill = y))+
   geom_boxplot()+
  # scale_x_log10()+
   coord_flip()
```

## Relationship between Pdays and the Outcome.
```{r}
bank_df %>%
   select(pdays,y) %>%
   ggplot(aes(pdays, fill = y))+
   geom_boxplot()+
   scale_x_log10()+
   coord_flip()
```

## Relationship between Previous and the Outcome.
```{r}
bank_df %>%
   select(previous,y) %>%
   ggplot(aes(previous, fill = y))+
   geom_boxplot()+
   scale_x_log10()+
   coord_flip()
```

## Relationship between poutcome and the Outcome.
```{r}
## alot of missing values, i think will remove this variable

p_rate(bank_df,poutcome)
```

## Some thoughts and Feature Engineering
Somme of the strong indicators here are that 
<ul>
    <li> Older People are more likely to subscribe </li>
    <li> Retired people and students have a high positivity rate compared to others </li>
    <li> The Divorced show a high rate too, but should that really be considered? </li>
    <li> People that earn above 10,000 Euros yearly are less likely to subscribe </li>
    <li> People that had a call duration of 8 minutes and higher were more likely to subscribe </li>
    <li> People that were going to subscribe didn't have to recieve many campaigns </li>
    
</ul>
Lets create a new variable containing most of our strong variables and test on the outcome

```{r}
## By creating a check list to get people who are either students or retired 
## who earn less than 10,000 Euros yearly and recieved a call that lasted for more than 8 Minutes, 
## the postivity rate is 47.83% from a total of 46 people.
bank_df %>%
   mutate(duration = duration/60,
          test = ifelse(job %in% c("student","retired") & balance < 10000 & duration > 8,
                        "They have the strong Variables","They dont have it")) %>%
   p_rate(test)
```

Well, that is all for now. I was able to really understand the dataset and when ML algorithms are applied i can get a better understanding. 



```{r}
bank_df %>%
   mutate(education = as.character(education),
          job = as.character(job),
          job = ifelse(job == "unknown",NA,job),
          education = ifelse(education == "unknown",NA,education)) %>%
   write_csv("bank_ML.csv")
   
## For Contact we will create a new variable to test for unknown
```
